Homicide Increases in the COVID-19 Era

### 2025-02-27 Draft1.qmd
cleaned up beginning data, moved messier raw data wrangling to the end. Cleaned data csvs are in the cleaned_data folder in Drive.

```{r}
crimes_with_zips <- read.csv("crimes_with_zips.csv") # just inner_join with the crime_la dataset
```

```{r}
# Secondary dataset - Median Income by ZCTA
california_median_income_2018_2023 <- list()
yrs <- 2018:2023
for (i in 1:length(yrs)) {
  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
  obj <- obj[-1,]
  obj$year <- yrs[[i]]
  california_median_income_2018_2023[[i]] <- obj
}

# adding a variable that's just the ZCTA zip code
cali_median_income$ZCTA <- stringr::str_sub(cali_median_income$NAME, -5)
```

```{r}
# Secondary dataset - COVID cases in Los Angeles County
covid_14day <- read.csv("Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")

@@

```{r, eval=FALSE}
library(sf) # for map
library(tidyverse) # using only for joining datasets to build map visual
library(RColorBrewer) # for map color scheme
library(tidygeocoder) # to get zip codes of addresses

@@

# Secondary datasets
california_median_income_2018_2023 <- list()
yrs <- 2018:2023
for (i in 1:length(yrs)) {
  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
  obj <- obj[-1,]
  obj$year <- yrs[[i]]
  california_median_income_2018_2023[[i]] <- obj
}

@@

### Zip Codes for Crimes {.unnumbered}
```{r, eval=FALSE}
library(sf)
library(dplyr)

@@

```{r}

@@

```{r, eval=FALSE}

@@

```{r, eval=FALSE}
crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
                                  format = "%m/%d/%Y %I:%M:%S %p", 
                                  tz = "America/Los_Angeles")

@@

```{r, eval=FALSE}
# Convert Date.Rptd to Date format (ignore time part)
#crime_la$Date.Rptd
crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")

@@

```{r, eval=FALSE}
# Remove missing values and empty strings in Vict.Sex column
crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))

@@

```{r, eval=FALSE}
zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

```{r, eval=FALSE}
# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)

@@

```{r, eval=FALSE}
# using latitude and longitude to obtain the zip codes of the crime locations
# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))

@@

```{r, eval=FALSE}
library(broom)

@@

```{r, eval=FALSE}
lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)

>>>>

# primary dataset
crime_la <- read.csv("Cleaned_Data/all_crimes_2010_present.csv")

# primary dataset zip code and geometry data
crimes_with_zips <- read.csv("Cleaned_Data/crimes_with_zips.csv") # just inner_join with the crime_la dataset

# Secondary dataset - Median Income by ZCTA
california_median_income_2018_2023 <- list()
yrs <- 2018:2023
for (i in 1:length(yrs)) {
  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
  obj <- obj[-1,]
  obj$year <- yrs[[i]]
  california_median_income_2018_2023[[i]] <- obj
}

cali_median_income <- read.csv("Cleaned_Data/cali_median_income.csv")

# Secondary dataset - COVID cases in Los Angeles County
covid_14day <- read.csv("Raw_Data/LA_County_Covid19_CSA_14day_case_death_table.csv")

@@

```{r Wrangling Main, eval=FALSE}
library(sf) # for map
library(tidyverse) # using only for joining datasets to build map visual
library(RColorBrewer) # for map color scheme
library(tidygeocoder) # to get zip codes of addresses

@@

```{r Combining Median Income, eval=FALSE}
# Secondary datasets
california_median_income_2018_2023 <- list()
yrs <- 2018:2023
for (i in 1:length(yrs)) {
  obj <- read.csv(paste0("Raw_Data/ACSST5Y",yrs[[i]],".S1903-Data.csv"), skip = 0)
  obj <- obj[-1,]
  obj$year <- yrs[[i]]
  california_median_income_2018_2023[[i]] <- obj
}

@@

### Zip Codes for Crimes {.unnumbered}
```{r Get Zip Codes, eval=FALSE}
library(sf)
library(dplyr)

@@

```{r Get Zip Codes 2, eval=FALSE}

@@

```{r Figure 1 Code, eval=FALSE}

@@

```{r Figure 2 Code, eval=FALSE}
crime_la$DateParsed <- as.POSIXct(crime_la$Date.Rptd, 
                                  format = "%m/%d/%Y %I:%M:%S %p", 
                                  tz = "America/Los_Angeles")

@@

```{r Figure 3 Code, eval=FALSE}
# Convert Date.Rptd to Date format (ignore time part)
#crime_la$Date.Rptd
crime_la$Date <- as.Date(substr(crime_la$Date.Rptd, 1, 10), format = "%m/%d/%Y")

@@

```{r Figure 4 Code, eval=FALSE}
# Remove missing values and empty strings in Vict.Sex column
crime_la <- subset(crime_la, Vict.Sex != "" & !is.na(Vict.Sex))

@@

```{r Figure 5 Wrangling, eval=FALSE}
zcta <- read_sf(dsn = "Raw_Data/cb_2020_us_zcta520_500k", layer = "cb_2020_us_zcta520_500k")
LA_zctas <- read.csv("Raw_Data/zctas.csv") #obtained using zipcodeR::search_city("Los Angeles", "CA") and then checking which zip codes are also ZCTAs

@@

```{r Figure 5 Code, eval=FALSE}
# quantiles for the entire dataset (from 2018 to 2023) to divide the median income into groups
qntls <- round(quantile(zcta_no_NAs$S1903_C03_001E, probs = seq(0, 1, length.out = 7)), 0)

@@

```{r Figure 6 Code, eval=FALSE}
# using latitude and longitude to obtain the zip codes of the crime locations
# crime_la$year <- as.integer(stringr::str_sub(crime_la$YearWeek, 1, 4))
# id_lat_long_crime_la <- subset(crime_la, select = c("LAT", "LON", "DR_NO", "year"))

@@

```{r Regression 1, eval=FALSE}
library(broom)

@@

```{r Regression 2, eval=FALSE}
lm_income_crimes <- lm(number_of_crimes ~ cali_median_income, data = crime_income_per_zip)
